---
title: "R log"
author: "Wu Yitong"
date: "2024-02-14"
output: github_document
---


# inverse matrix

```{r matrix 1}
m <- matrix(data=rnorm(9), nrow=3)
m
```

Let's try `-1` first.

```{r matrix 2}
m^(-1)
```
This is right way.

```{r matrix 3}
solve(m)
```


# ifelse

So in this case, `ifelse` can replace `if` + `else`.
```{r ifelse 1}
ifelse(TRUE, 1, 0)
if (TRUE) 1 else 0
```

But for a list, `ifelse` can only return the first value.
```{r ifelse 2}
ifelse(TRUE, c(1, 2, 3), c(5, 6, 7))
if (TRUE) c(1, 2, 3) else c(5, 6, 7)
```


# [a:a+1]

Be careful about the `[]`.
```{r [a:a+1] 1}
l <- c(0.1, 0.2, 0.3, 0.4, 0.5)

l[1:2 + 1] # == l[2:3]
l[1 + 1:2] # == l[2:3]
l[1 + 1:2 + 1] # == l[3:4]

l[1:(2 + 1)] # == l[1:3]
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# match.call()

We saw lots of `match.call()` in many source code, including `glm`, `rstanarm`, etc.

Here is an example from `glm`,
```{r glm, eval=FALSE}
glm <- function(formula, family = gaussian, data, weights,
                subset, na.action, start = NULL,
                etastart, mustart, offset,
                control = list(...),
                model = TRUE, method = "glm.fit",
                x = FALSE, y = TRUE,
                contrasts = NULL, ...) {
  call <- match.call()

  ...

  mf <- match.call(expand.dots = FALSE)
  m <- match(c(
    "formula", "data", "subset", "weights", "na.action",
    "etastart", "mustart", "offset"
  ), names(mf), 0L)
  mf <- mf[c(1L, m)]
  mf$drop.unused.levels <- TRUE
  ## need stats:: for non-standard evaluation
  mf[[1L]] <- quote(stats::model.frame)
  mf <- eval(mf, parent.frame())
  if (identical(method, "model.frame")) return(mf)

  ...

```

and `rstanarm`.
```{r rstanarm, eval=FALSE}
stan_glm <-
  function(formula,
           family = gaussian(),
           data,
           weights,
           subset,
           na.action = NULL,
           offset = NULL,
           model = TRUE,
           x = FALSE,
           y = TRUE,
           contrasts = NULL,
           ...,
           prior = default_prior_coef(family),
           prior_intercept = default_prior_intercept(family),
           prior_aux = exponential(autoscale = TRUE),
           prior_PD = FALSE,
           algorithm = c("sampling", "optimizing", "meanfield", "fullrank"),
           mean_PPD = algorithm != "optimizing" && !prior_PD,
           adapt_delta = NULL,
           QR = FALSE,
           sparse = FALSE) {
    algorithm <- match.arg(algorithm)
    family <- validate_family(family)
    validate_glm_formula(formula)
    data <- validate_data(data, if_missing = environment(formula))

    call <- match.call(expand.dots = TRUE)
    mf <- match.call(expand.dots = FALSE)
    m <- match(c("formula", "subset", "weights", "na.action", "offset"),
      table = names(mf), nomatch = 0L
    )
    mf <- mf[c(1L, m)]
    mf$data <- data
    mf$drop.unused.levels <- TRUE
    mf[[1L]] <- as.name("model.frame")
    mf <- eval(mf, parent.frame())
    mf <- check_constant_vars(mf)
    mt <- attr(mf, "terms")
    Y <- array1D_check(model.response(mf, type = "any"))
    if (is.empty.model(mt)) {
      stop("No intercept or predictors specified.", call. = FALSE)
    }
    X <- model.matrix(mt, mf, contrasts)
    
    ...
    
```

This is definition of `match.call()`.
```{r mc def, eval=FALSE}
match.call(
  definition = sys.function(sys.parent()),
  call = sys.call(sys.parent()),
  expand.dots = TRUE,
  envir = parent.frame(2L)
)
```

Lets see what `match.call()` does.
```{r mc usage 1.1}
func_1 <- function(a, b, c, ...) {
  return(match.call())
}

func_1(1, 2, 3)
```

What if we have extra augments.
```{r mc usage 1.2}
func_1(1, 2, 3, 4, 5)
```

As you can see, `match.call()` can help us collect the augments passing through the current function.

Lets use another function.
```{r multi-mc  2.1}
# multiply a and b only
multiply_inner <- function(a, b, c, ...) {
  if (is.null(c)) cat("value c is not available\n") 
  else cat("the value of c is:", c, "\n")
  return(a*b)
}
```

Put the similar `match.call` code from above into this outer function.

And lets see what changed during the running.
```{r multi-mc 2.2}
multiply_outer_1 <- function(a, b, c, d, ...) {
  cat("the value of d is:", d, "\n")
  
  ## expand.dot = TRUE => each extra value is a single value
  call <- match.call() # expand.dots = TRUE by defaut
  # [1] "call: multiply_outer"
  # [2] "call: 2"             
  # [3] "call: 3" 
  # [4] "call: 4"             
  # [5] "call: 5"   
  # [6] "call: 6"  
  # [6] "call: 7"  
  
  ## expand.dot = FALSE => all yje extra value will store as a whole one
  mf <- match.call(expand.dots = FALSE)
  # [1] "mf: multiply_outer"
  # [2] "mf: 2"             
  # [3] "mf: 3"   
  # [3] "mf: 4"   
  # [3] "mf: 5"   
  # [4] "mf: pairlist(6, 7)" 
  
  ## select augments a and b only
  m <- match(c("a", "b", "c"), names(mf), 0L)
  # [1] "m: 2" "m: 3" "m: 4"
  
  ## 1L is the current function
  mf <- mf[c(1L, m)] 
  # [1] "mf: multiply_outer" 
  # [2] "mf: 2"             
  # [3] "mf: 3"   
  # [3] "mf: 4"   
  
  ## change 1L into other function (here is multiply_inner)
  mf[[1L]] <- quote(multiply_inner)
  # [1] "mf: multiply_inner" 
  # [2] "mf: 2"             
  # [3] "mf: 3"  
  # [3] "mf: 4"  
  
  ## run the inner function we just assigned
  mf <- eval(mf, parent.frame())
  # [1] "mf: 6"
  
  return(mf)
}

multiply_outer_1(2,3,4,5,6,7)
```

So, you may ask, why we cannot just use...
```{r simply usage}
multiply_outer_2 <- function(a, b, c, d, ...) {
  cat("the value of d is:", d, "\n")
  return(multiply_inner(a, b, c))
}

multiply_outer_2(2,3,4,5,6)
```
The answer is YES, in this case, using above code would suddenly simplify the process.

But, what if you want more flexibility.
```{r}
some_glm <- function(formula, model, ...) {
  
  call <- match.call()
  # call: customize_summary(formula = mpg ~ cyl, model = lm_model, method = "bootstrap", mean = TRUE, median = TRUE)
  
  args <- as.list(call)[-1]  # Remove the function name
  # args:
  # $formula
  # y ~ x1+x2
  # 
  # $model
  # lm_model
  # 
  # $method
  # [1] "bootstrap"
  # 
  # $mean
  # [1] TRUE
  # 
  # $median
  # [1] TRUE
  #
  # $sd
  # [1] FALSE

  ## Extract the method argument
  method_arg <- args$method
  # [1] method_arg: bootstrap
  
  ## Extract other relevant arguments
  statistics <- c("mean", "median", "sd")
  selected_stats <- intersect(names(args)[sapply(args, function(x) identical(x, TRUE))], statistics)

  return(selected_stats)
}

# Example usage
some_glm(y ~ x1+x2, lm_model, method = "bootstrap", mean = TRUE, median = TRUE, sd=FALSE)
```
So you can play with the augments!

